```{r global_options, echo=FALSE}
opts_chunk$set(dev="svg", fig.width=8, fig.height=6, echo=FALSE, warning=FALSE, message=FALSE);
```

```{r preamble}
library(io);
library(mmalign);
library(viridis);
library(reshape2);
library(ggplot2);
library(party);
library(RColorBrewer);

source("plot.R");
```

```{r preprocess, cache=TRUE}
d <- qread("data.rds");
e <- qread("data_expanded.rds");

xd <- d[, -ncol(d)];
xe <- e[, -ncol(e)];

y <- d$Cover_Type;
yd <- d[, "Cover_Type", drop=FALSE];

cover.colours <- brewer.pal(7, "Set1");

xet <- t(as.matrix(xe));

p <- pca(xet, center.features=TRUE, scale.features=TRUE);

not.outlier <- apply(p$Z, 2, function(x) all(abs(x) < 10));

set.seed(1234);
sub.idx <- sample(which(not.outlier), 5000);

xes <- xe[sub.idx, ];
xds <- xd[sub.idx, ];
xets <- xet[, sub.idx];
yds <- yd[sub.idx, , drop=FALSE];
ds <- d[sub.idx, ];
zs <- p$Z[, sub.idx];
```

## Principal component analysis

We first examine whether the features are globally correlated with the
response variable, forest cover type. Outliers have been removed from PCA
plots to avoid excessive skew.

```{r pca}
mmalign:::pca_plot_base(zs, yds, aes(colour=Cover_Type), vars=p$var, dims=1:2) +
	geom_point(alpha=0.3) +
	theme(plot.title=element_text(hjust=0)) +
	scale_colour_manual(values=cover.colours) +
	ggtitle("Principal component analysis of feature data")

mmalign:::pca_plot_base(zs, yds, aes(colour=Cover_Type), vars=p$var, dims=2:3) +
	geom_point(alpha=0.3) + 
	theme(plot.title=element_text(hjust=0)) +
	scale_colour_manual(values=cover.colours) +
	ggtitle("Principal component analysis of feature data")

mmalign:::pca_plot_base(zs, yds, aes(colour=Cover_Type), vars=p$var, dims=3:4) +
	geom_point(alpha=0.3) + 
	theme(plot.title=element_text(hjust=0)) +
	scale_colour_manual(values=cover.colours) +
	ggtitle("Principal component analysis of feature data")

mmalign:::pca_plot_base(zs, yds, aes(colour=Cover_Type), vars=p$var, dims=4:5) +
	geom_point(alpha=0.3) + 
	theme(plot.title=element_text(hjust=0)) +
	scale_colour_manual(values=cover.colours) +
	ggtitle("Principal component analysis of feature data")
```


## Correlative features

To identify features that correlate with the forest cover type, we correlate
each feature with each forest cover type.

```{r cors, fig.height=4}
ye <- lapply(levels(y),
	function(v) {
		as.numeric(y == v)
	}
);
names(ye) <- levels(y);
ye <- data.frame(ye, check.names=FALSE);

cors <- apply(xe, 2, function(x) cor(x, ye));
rownames(cors) <- colnames(ye);

cors.dm <- melt(cors, varnames=c("y", "x"));

g <- ggplot(cors.dm, aes(x=x, y=y, fill=value)) + 
	geom_tile(color="white", size=0.1) +
	scale_fill_viridis(name="correlation") +
	coord_equal() + 
	labs(x=NULL, y=NULL, title="Correlations between forest features and cover types") +
	theme(plot.title=element_text(hjust=0)) +
	theme(axis.ticks=element_blank()) +
	theme(axis.text.y=element_text(size=7)) +
	theme(axis.text.x=element_text(size=7, angle=45, hjust=1)) +
	theme(legend.title=element_text(size=8)) +
	theme(legend.text=element_text(size=6))
g
```

The heatmap reveals that elevation, some wilderness areas, and some
soil types are correlated with some forest cover types.


```{r}
f <- with(ds,
	Cover_Type ~ Elevation + Aspect + Slope + Soil_Type + Wilderness_Area +
		Horizontal_Distance_To_Hydrology + Vertical_Distance_To_Hydrology + 
		Horizontal_Distance_To_Roadways + Hillshade_9am + Hillshade_Noon +
		Hillshade_3pm + Horizontal_Distance_To_Fire_Points
);
```

## Conditional inference trees

Finally, we illustrate how each candidate feature can help predict forest cover type
using decision trees.

```{r tree_elevation, fig.height=8}
ct.e <- ctree(Cover_Type ~ Elevation, data = ds, controls=ctree_control(maxdepth=2));
plot(ct.e, main="Cover_Type ~ Elevation",
	tp_args=list(fill=cover.colours, col=NA, gap=0),
	terminal_panel=node_barplot_rot);
```

```{r tree_wilderness, fig.height=8}
ct.w <- ctree(Cover_Type ~ Wilderness_Area, data = ds);
plot(ct.w, main="Cover_Type ~ Wilderness_Area",
	tp_args=list(fill=cover.colours, col=NA, gap=0),
	terminal_panel=node_barplot_rot);
```

```{r tree_soil_type, fig.height=8}
ct.s <- ctree(Cover_Type ~ Soil_Type, data = ds,
							controls=ctree_control(maxdepth=2));
plot(ct.s, main="Cover_Type ~ Soil_Type",
	tp_args=list(fill=cover.colours, col=NA, gap=0),
	terminal_panel=node_barplot_rot);
```

